### 2024-06-06 manuscript/Manuscript.tex
COS course day 2 updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Arial}
  \setmonofont[]{DejaVu Sans Mono}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[lmargin=0.8in,rmargin=0.8in,tmargin=0.8in,bmargin=0.8in]{geometry}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\usepackage{lineno}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={New paper template},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{New paper template}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\raggedright

Author 1\textsuperscript{1,2,*} , Author 2\textsuperscript{1,2}, Author
3\textsuperscript{3}, Author 4E\textsuperscript{4}
\newline\textsuperscript{1}Heidelberg University, Centre for Organismal
Studies (COS), 69120 Heidelberg, Germany
\newline\textsuperscript{2}Affiliation 2
\newline\textsuperscript{3}Affiliation 3
\newline\textsuperscript{4}Affiliation 4\newline
\textsuperscript{*}Correspondence: xy@cos.uni-heidelberg.de}
\author{}
\date{}

\begin{document}
\maketitle

\linenumbers

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\setstretch{1.2}
\section{Abstract}\label{abstract}

\ul{This} is an R \textbf{Markdown} \emph{document.}
\textsuperscript{Markdown} is a simple formatting syntax for authoring
HTML, PDF, and MS Word documents. For more details on using R Markdown
see \url{http://rmarkdown.rstudio.com}.

\textbf{text in bold} \emph{italic} \ul{underline}

\section{Introduction}\label{introduction}

You can add references either by referring to their id in the .bib file
e.g., (Marinković et al., 2019), or by switching to the visual editor
(Cogwheel in the .Rmd menu -\textgreater{} Use Visual Editor). {[}Jokura
et al. (2023){]}(Jokura et al., 2023)(Jacobs and Ryu, 2023)

It is now a test to edit the text and see how the changes show up on
GitHub.

Test of git show.

In the visual editor mode, go to `Insert' -\textgreater{} @ Citation

You can select a Zotero library, PubMed, CrossRef etc. and insert the
citations. (Jacobs and Ryu, 2023)

The easiest way is to use the command line:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{curl} \AttributeTok{{-}LH} \StringTok{"Accept: application/x{-}bibtex"}\NormalTok{ https://doi.org/10.7554/eLife.91258.1 }\OperatorTok{\textgreater{}\textgreater{}}\NormalTok{ references.bib}
\end{Highlighting}
\end{Shaded}

\emph{Platynereis dumerilii} is a marine annelid\ldots{} (Ozpolat et
al., 2021)

The references are stored in manuscript/references.bib (need to be
defined in the Yaml header). This file will automatically updated when
you insert a new reference through the Visual editor \textgreater{}
Insert \textgreater{} Citations.

{[}Ozpolat et al. (2021){]}(Lohmann et al., 2023; Saile et al., 2023; Wu
et al., 2020)(Lohmann et al., 2001)

In this documents, references will be formatted in the style of eLife.
This is defined in the Yaml header under: csl: elife.csl. The elife.csl
file is saved in the /manuscript folder.

If you would like to use a different citation format, download the
respective .csl file (e.g., from the Zotero style repository
\url{https://www.zotero.org/styles}), save it in the /manuscript folder
of the project and change the Yaml to csl: your\_favourite\_journal.csl.

\subsection{Results}\label{results}

Inserting Figures

Test comment here.

You can add your figures into the rendered document. We saved the
figures into /manuscript/figures or /manuscript/figure\_supplements and
can insert them from there. We use knitr::include\_graphics for this.
The title and legend can also be edited, as will as the width of the
output figure. Test comment behaviour:

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/Figure1.png}

}

\caption{\textbf{Figure 1. A figure} (A) A nice picture. (B) legend. (C)
(D)}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/Figure_complex.png}

}

\caption{\textbf{Figure 1. Our nice figure from yesterday} (A) A nice
picture. (B) legend. (C) (D)}

\end{figure}%

\subsection{Equations}\label{equations}

Equations can also be inserted, Insert -\textgreater{} Display Math:

\[
\bar{X} = \frac{\sum_{i=1}^{n} x_{i}}{n}
\]

\hfill\break

\subsection{Sourcing code and working with
variable}\label{sourcing-code-and-working-with-variable}

The mean value of Nanog expression was 0.0909 indicating that Nanog is
downregulated. The `analysis/scripts/statistics\_for\_paper.R' script is
sourced and it runs but the output is not included in the knitted
output. But we can access the variables defined in the sourced script
simply by adding ` r var\_name ` between ` backticks, in this case
max\_PRC value is 21 (now this number comes from our sourced script).

If we update the data, the script can recalculate the variable we want
to refer to in the text and update the number.

\subsection{Acknowledgements}\label{acknowledgements}

We would like to thank the Jekely lab for the R project template
(\url{https://github.com/JekelyLab/new_paper_template}) we used to write
this paper. This work was funded by \ldots{}

\subsection{Materials and Methods}\label{materials-and-methods}

You can insert tables from source data, such as .csv or Excel files and
render them in html with the tinytable package.

Alternatively, you can use the Markdown grid table format. For more
complex tables, you can use the
\href{https://www.tablesgenerator.com/markdown_tables}{tablesgenerator}
online grid table editor/converter (e.g.~converts csv or excel files).

The output may differ between html and pdf, for most consistent results
use the grid table format described
\href{https://quarto.org/docs/authoring/tables.html}{here}.

\textbf{Key Resources Table}

\begin{verbatim}
Warning: The `placement` argument in `tt()` is deprecated. Please use this
instead: `theme_tt(table, 'placement')`
\end{verbatim}

\begin{table}[H]
\centering
\begin{tblr}[         %% tabularray outer open
]                     %% tabularray outer close
{                     %% tabularray inner open
width={1\linewidth},
colspec={X[]X[]X[]X[]X[]},
hlines, vlines,
}                     %% tabularray inner close
Reagent type (species) or resource & Designation & Source or reference & Identifiers & Additional information \\
biological sample (N. vectensis)     & larval, juvenile and adult N. vectensis        & Specimens obtained form the Marine Invertebrate Culture Unit of the University of Exeter & N/A      & NA                                                                                                               \\
biological sample (cDNA)             & cDNA obtained from N. vectensis                & this study                                                                               & N/A      & RNA extracted with Trizol and cDNA synthesized with cDNA synthesis kit according to manufacturers recommendation \\
biological sample (peptide extract)  & peptide extracts obtained from N. vectensis    & this study                                                                               & N/A      & Peptides extracted from N. vectensis according to protocol explained in Material and Methods                     \\
genetic reagent (cDNA synthesis)     & SuperScript™ III First-Strand Synthesis System & Invitrogen (from ThermoFisher)                                                           & 18080051 & NA                                                                                                               \\
genetic reagent (Polymerase)         & Q5® Hot Start High-Fidelity DNA Polymerase     & New England Biolabs                                                                      & M0493L   & NA                                                                                                               \\
genetic reagent (DNA assembly)       & NEBuilder® HiFi DNA Assembly Master Mix        & New England Biolabs                                                                      & E2621L   & NA                                                                                                               \\
genetic reagent (restriction enzyme) & EcoRV restriction enzyme                       & New England Biolabs                                                                      & R3195L   & NA                                                                                                               \\
genetic reagent (restriction enzyme) & Afl2 restriction enzyme                        & New England Biolabs                                                                      & R0520L   & NA                                                                                                               \\
\end{tblr}
\end{table}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3000}}@{}}
\caption{Grid Table example}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Col1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col4
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col5
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Col1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col4
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Col5
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
a & b & c & d & e \\
d & & & & \\
\end{longtable}

\subsection{Complex grid table
example}\label{complex-grid-table-example}

This table was generated by tt() as the output of an r chunk in a Quarto
doc. For larger multi-page tables, this method gives correct page breaks
in the pdf and html outputs. You can change the relative column widths
with \{tbl-colwidths=``{[}10,20,20,20,30{]}''\} placed after the table
caption declaration at the end.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3000}}@{}}
\caption{More complex Grid Table example}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Reagent type (species) or resource
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Designation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source or reference
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Identifiers
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Additional information
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Reagent type (species) or resource
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Designation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source or reference
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Identifiers
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Additional information
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
biological sample (N. vectensis) & larval, juvenile and adult N.
vectensis & Specimens obtained form the Marine Invertebrate Culture Unit
of the University of Exeter & N/A & NA \\
biological sample (cDNA) & cDNA obtained from N. vectensis & this study
& N/A & RNA extracted with Trizol and cDNA synthesized with cDNA
synthesis kit according to manufacturers recommendation \\
biological sample (peptide extract) & peptide extracts obtained from N.
vectensis & this study & N/A & Peptides extracted from N. vectensis
according to protocol explained in Material and Methods \\
genetic reagent (cDNA synthesis) & SuperScript™ III First-Strand
Synthesis System & Invitrogen (from ThermoFisher) & 18080051 & NA \\
genetic reagent (Polymerase) & Q5® Hot Start High-Fidelity DNA
Polymerase & New England Biolabs & M0493L & NA \\
genetic reagent (DNA assembly) & NEBuilder® HiFi DNA Assembly Master Mix
& New England Biolabs & E2621L & NA \\
genetic reagent (restriction enzyme) & EcoRV restriction enzyme & New
England Biolabs & R3195L & NA \\
genetic reagent (restriction enzyme) & Afl2 restriction enzyme & New
England Biolabs & R0520L & NA \\
\end{longtable}

\subsection{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Jacobs2023}
Jacobs EAK, Ryu S. 2023. Larval zebrafish as a model for studying
individual variability in translational neuroscience research.
\emph{Frontiers in Behavioral Neuroscience} \textbf{17}.
doi:\href{https://doi.org/10.3389/fnbeh.2023.1143391}{10.3389/fnbeh.2023.1143391}

\bibitem[\citeproctext]{ref-Jokura_2023}
Jokura K, Ueda N, Gühmann M, Yañez-Guerra LA, Słowiński P, Wedgwood KCA,
Jékely G. 2023. Nitric oxide feedback to ciliary photoreceptor cells
gates a UV avoidance circuit.
doi:\href{https://doi.org/10.7554/elife.91258.1}{10.7554/elife.91258.1}

\bibitem[\citeproctext]{ref-Lohmann2023}
Lohmann J, de Luxán-Hernández C, Gao Y, Zoschke R, Weingartner M. 2023.
Arabidopsis translation factor eEF1Bγ impacts plant development and is
associated with heat-induced cytoplasmic foci. \emph{Journal of
Experimental Botany} \textbf{74}:2585--2602.
doi:\href{https://doi.org/10.1093/jxb/erad050}{10.1093/jxb/erad050}

\bibitem[\citeproctext]{ref-lohmann2001}
Lohmann JU, Hong RL, Hobe M, Busch MA, Parcy F, Simon R, Weigel D. 2001.
A Molecular Link between Stem Cell Regulation and Floral Patterning in
Arabidopsis. \emph{Cell} \textbf{105}:793--803.
doi:\href{https://doi.org/10.1016/s0092-8674(01)00384-1}{10.1016/s0092-8674(01)00384-1}

\bibitem[\citeproctext]{ref-Marinkovi2020}
Marinković M, Berger J, Jékely G. 2019. Neuronal coordination of motile
cilia in locomotion and feeding. \emph{Philosophical Transactions of the
Royal Society B: Biological Sciences} \textbf{375}:20190165.
doi:\href{https://doi.org/10.1098/rstb.2019.0165}{10.1098/rstb.2019.0165}

\bibitem[\citeproctext]{ref-ozpolat2021}
Ozpolat BD, Randel N, Williams EA, Bezares-Calderón LA, Andreatta G,
Balavoine G, Bertucci PY, Ferrier DEK, Gambi MC, Gazave E,
Handberg-Thorsager M, Hardege J, Hird C, Hsieh Y-W, Hui J, Mutemi KN,
Schneider SQ, Simakov O, Vergara HM, Vervoort M, Jékely G,
Tessmar-Raible K, Raible F, Arendt D. 2021. The Nereid on the rise:
\emph{Platynereis} as a model system. \emph{Zenodo}.
doi:\href{https://doi.org/10.5281/ZENODO.4907400}{10.5281/ZENODO.4907400}

\bibitem[\citeproctext]{ref-Saile2023}
Saile J, Wießner-Kroh T, Erbstein K, Obermüller DM, Pfeiffer A, Janocha
D, Lohmann J, Wachter A. 2023. SNF1-RELATED KINASE 1 and TARGET OF
RAPAMYCIN control light-responsive splicing events and developmental
characteristics in etiolated Arabidopsis seedlings. \emph{The Plant
Cell} \textbf{35}:3413--3428.
doi:\href{https://doi.org/10.1093/plcell/koad168}{10.1093/plcell/koad168}

\bibitem[\citeproctext]{ref-Wu2020}
Wu H, Qu X, Dong Z, Luo L, Shao C, Forner J, Lohmann JU, Su M, Xu M, Liu
X, Zhu L, Zeng J, Liu S, Tian Z, Zhao Z. 2020. WUSCHEL triggers innate
antiviral immunity in plant stem cells. \emph{Science}
\textbf{370}:227--231.
doi:\href{https://doi.org/10.1126/science.abb7360}{10.1126/science.abb7360}

\end{CSLReferences}

\end{document}