### 2024-06-06 manuscript/table_test.qmd

---
title: "table_test"
format: html
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Table

|   |   |   |   |   |
|---|---|---|---|---|
|   |   |   |   |   |
|   |   |   |   |   |
|   |   |   |   |   |

| Reagent type (species) or resource         | Designation                                        | Source or reference                                                                                | Identifiers          | Additional information                                                                                           |
|--------------------------------------------|----------------------------------------------------|----------------------------------------------------------------------------------------------------|----------------------|------------------------------------------------------------------------------------------------------------------|
|                                            |                                                    |                                                                                                    |                      |                                                                                                                  |
| biological sample (N. vectensis)           | larval, juvenile and adult N. vectensis            | Specimens obtained form the Marine Invertebrate Culture Unit of the University of Exeter           | N/A                  |                                                                                                                  |
| biological sample (cDNA)                   | cDNA obtained from N. vectensis                    | this study                                                                                         | N/A                  | RNA extracted with Trizol and cDNA synthesized with cDNA synthesis kit according to manufacturers recommendation |
| biological sample (peptide extract)        | peptide extracts obtained from N. vectensis        | this study                                                                                         | N/A                  | Peptides extracted from N. vectensis according to protocol explained in Material and Methods                     |
| genetic reagent (cDNA synthesis)           | SuperScript™ III First-Strand Synthesis System     | Invitrogen (from ThermoFisher)                                                                     | 18080051             |                                                                                                                  |
| genetic reagent (Polymerase)               | Q5® Hot Start High-Fidelity DNA Polymerase         | New England Biolabs                                                                                | M0493L               |                                                                                                                  |
| genetic reagent (DNA assembly)             | NEBuilder® HiFi DNA Assembly Master Mix            | New England Biolabs                                                                                | E2621L               |                                                                                                                  |
| genetic reagent (restriction enzyme)       | EcoRV restriction enzyme                           | New England Biolabs                                                                                | R3195L               |                                                                                                                  |
| genetic reagent (restriction enzyme)       | Afl2 restriction enzyme                            | New England Biolabs                                                                                | R0520L               |                                                                                                                  |
| genetic reagent (restriction enzyme)       | Hind3 restriction enzyme                           | New England Biolabs                                                                                | R3104L               |                                                                                                                  |
| genetic reagent (restriction enzyme)       | BamH1 restriction enzyme                           | New England Biolabs                                                                                | R3136L               |                                                                                                                  |
| genetic reagent (restriction enzyme)       | EcoRI restriction enzyme                           | New England Biolabs                                                                                | R3101L               |                                                                                                                  |
| recombinant DNA reagent (plasmid)          | pcDNA3.1(+)                                        | Invitrogen (from ThermoFisher)                                                                     | V79020               |                                                                                                                  |
| recombinant DNA reagent (plasmid)          | pRK5-Gqi9                                          | AddGene                                                                                            | 125711               |                                                                                                                  |
| recombinant DNA reagent (plasmid)          | pcDNA3.1(+)-G5A                                    | this study                                                                                         | N/A                  | sequence information provided in Supplementary file 7                                                            |
| recombinant DNA reagent (plasmid)          | pcDNA3.1(+)-Gqi5/9                                 | this study                                                                                         | N/A                  | sequence information provided in Supplementary file 7                                                            |
| recombinant DNA reagent                    | cloned N. vectensis GPCRs in pcDNA3.1(+)           | identified in this study                                                                           | N/A                  | full list of cloned GPCRs with sequences in Supplementary file 8                                                 |
| recombinant DNA reagent                    | synthesized N. vectensis GPCRs in pcDNA3.1(+)      | identified in this study, codon optimized and synthesized by GenScript                             | N/A                  | full list of synthesized GPCRs with original and codon optimized sequences in Supplementary file 8               |
| sequence-based reagent                     | cloning primers to create Gqi5/9                   | this study                                                                                         | N/A                  | sequence information provided in Supplementary file 7                                                            |
| sequence-based reagent                     | GPCR cloning primers                               | this study                                                                                         | N/A                  | full list of primers with sequences in Supplementary file 8                                                      |
| peptide, recombinant protein               | custom peptides                                    | identified in this study, synthesized by GenScript                                                 | N/A                  | full list of peptides with sequences and purity in Supplementary file 2 and 3                                    |
| commercial assay or kit (PCR clean up)     | Monarch® PCR & DNA Cleanup Kit (5 μg)              | New England Biolabs                                                                                | T1030L               |                                                                                                                  |
| commercial assay or kit (Miniprep)         | GeneJET Plasmid Miniprep Kit                       | Thermo Fisher Scientific                                                                           | K0503                |                                                                                                                  |
| strain, strain background (E. coli, TOP10) | One Shot™ TOP10 Chemically Competent E. coli       | Thermo Fisher Scientific                                                                           | C404010              | E. coli strain used for general cloning                                                                          |
| cell line (HEK293)                         | HEK293 Cells expressing GFP-AEQUORIN in Cytoplasma | Angio-Proteomie                                                                                    | cAP-0200GFP-AEQ-Cyto |                                                                                                                  |
| chemical compound, drug                    | DMEM                                               | Gibco (from ThermoFisher)                                                                          | 10566016             |                                                                                                                  |
| chemical compound, drug                    | OptiMEM                                            | Gibco (from ThermoFisher)                                                                          | 11058021             |                                                                                                                  |
| chemical compound, drug                    | FBS                                                | Gibco (from ThermoFisher)                                                                          | 10500064             |                                                                                                                  |
| chemical compound, drug                    | PEI (Polyethylenimine, 25k Mw)                     | Sigma-Aldrich                                                                                      | 408727               | used 0.3 µl of a 1 mg/ml stock solution per 100 ng DNA                                                           |
| chemical compound, drug                    | Transfectamine 5000                                | AAT Bioquest                                                                                       | 60022                | used 0.3 µl per 100 ng of DNA                                                                                    |
| chemical compound, drug                    | Coelenterazine-h                                   | Promega                                                                                            | S2011                | diluted to 2 mM in Ethanol and used at a final concentration of 4 µM in the assays                               |
| chemical compound, drug                    | TRIzol™ Reagent                                    | Invitrogen (from ThermoFisher)                                                                     | 15596026             |                                                                                                                  |
| other                                      | Corning® 96 Well White Polystyrene Microplate      | Corning (from Sigma-Aldrich)                                                                       | CLS3903              | cell culture treated assay plates                                                                                |
| other                                      | Nunc™ EasYFlask™ T75 Cell Culture Flasks           | Nunc (from ThermoFisher)                                                                           | 156499               | cell culture flasks                                                                                              |
| other                                      | Flexstation 3 Multimode Microplate Reader          | Molecular Devices                                                                                  |                      | plate reader                                                                                                     |
| software, algorithm                        | SoftMax®  Pro 7                                    | Molecular Devices                                                                                  |                      |                                                                                                                  |
| software, algorithm                        | R                                                  | https://cran.rstudio.com/                                                                          |                      |                                                                                                                  |
| software, algorithm                        | RStudio                                            | https://posit.co/download/rstudio-desktop/                                                         |                      |                                                                                                                  |
| software, algorithm                        | SignalP4.1                                         | https://services.healthtech.dtu.dk/services/SignalP-6.0/                                           |                      |                                                                                                                  |
| software, algorithm                        | NeuroPID                                           | https://bio.tools/neuropid                                                                         |                      |                                                                                                                  |
| software, algorithm                        | HMMER3.1b2                                         | http://hmmer.org/download.html                                                                     |                      |                                                                                                                  |
| software, algorithm                        | CD-HIT                                             | https://sites.google.com/view/cd-hit                                                               |                      |                                                                                                                  |
| software, algorithm                        | TransDecoder                                       | https://github.com/TransDecoder/TransDecoder                                                       |                      |                                                                                                                  |
| software, algorithm                        | CLANS (desktop version)                            | https://mybiosoftware.com/clans-20101007-visualize-protein-families-based-pairwise-similarity.html |                      |                                                                                                                  |
| software, algorithm                        | CLANS (online toolkit)                             | https://toolkit.tuebingen.mpg.de/tools/clans                                                       |                      |                                                                                                                  |
| software, algorithm                        | Phobius                                            | https://phobius.sbc.su.se/data.html                                                                |                      |                                                                                                                  |
| software, algorithm                        | muscle alignment tool                              | https://drive5.com/muscle5/                                                                        |                      |                                                                                                                  |
| software, algorithm                        | MAFFT v7                                           | https://mafft.cbrc.jp/alignment/software/                                                          |                      |                                                                                                                  |
| software, algorithm                        | trimAl                                             | http://trimal.cgenomics.org/trimal                                                                 |                      |                                                                                                                  |
| software, algorithm                        | IQ-tree2                                           | http://www.iqtree.org/                                                                             |                      |                                                                                                                  |
| software, algorithm                        | Fasttree                                           | http://www.microbesonline.org/fasttree/                                                            |                      |                                                                                                                  |

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).