### 2024-11-01 presentation/RStudio_paper_writing_course_presentation_part2.qmd

---
title: "Rstudio in Data Science and Paper Writing"
subtitle: "Part 2"
author: "Gáspár Jékely"
format:
  revealjs: 
    theme: [default, custom.scss]
    slide-number: false
    code-fold: false
    code-summary: "Code"
    chalkboard: 
      buttons: false
    preview-links: auto
scrollable: true
---

## Course program

<br>
- **Section 1**: Why open science? Project management, data import, tidy data
- **Section 2**: Plotting, figure assembly
- **Section 3**: Manuscript writing, tables, references, data and figure embedding
- **Section 4**: GitHub, collaboration, sharing, version control

```{r eval=TRUE, echo=FALSE}
source("../analysis/scripts/packages_and_functions.R")
```