### 2024-11-11 manuscript/Test_quarto.qmd

---
title: "R course report - group 1"
author: "name1, name2, name3"
format: html
---

## Quarto
### Heading 2
#### Heading 3

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
print('hi, this is an R command executed')
```

```{bash}
echo hi
```

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
#| eval: true
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Next section